---
title:  "Central Limit Theorem App" 
date: "April 2, 2019"
output: 
  ioslides_presentation:
    logo: logo.png
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

## Central Limit Theorem

In probability theory, the **central limit theorem** establishes that, in some situations, when independent random variables are added, their properly normalized sum tends toward a **normal distribution** even if the original variables themselves are not normally distributed. 

The theorem is a key concept in probability theory because it implies that probabilistic and statistical methods that work for normal distributions can be applicable to many problems involving other types of distributions. 


(source: Wikipedia)



## The App

- Our app demonstrates the central limit theorem **interactively** for the exponential distribution

<p style="text-align: center;">
PDF:  f(x;λ) = λ\*exp(-λ\*x)
</p>

- The app simulates the distribution of averages of random samples of exponential distributions, 
and outputs the corresponding histogram.

- The exponential distribution is simulated in R with ```rexp(n, lambda)``` where ```lambda``` is the rate
parameter and ```n``` the sample size.

- The histogram is produced using the ggplot2 data visualization package for R.



## User Input

The user can set:

- the value of the rate parameter λ 
- the sample size 

The user can choose to display over the histogram:

- the sample and theoretical means
- the probability density of the normal distribution 

<p style="text-align: center;">
N(1/λ,1/λ*sqrt(1/n))
</p>

to which the distribution of sample averages approximates



## Example of Output

```{r}
lambda <- 0.2 #Set rate to 0.2
sampleSize <- 40 #Set sample size to 40
set.seed(307) #Set seed for reproducibility

mns = NULL #Initiate variable to store the outcome of the simulations
for (i in 1 : 1000) mns = c(mns, mean(rexp(sampleSize,lambda)))

ggplot(data.frame(mns), aes(x=mns)) +
geom_histogram(aes(y=..density..), binwidth=0.2, fill='lemonchiffon2',
color="black", alpha = .8) +
labs(x="sample averages") +
stat_function(fun=dnorm, args=list(mean=5, sd=sqrt(0.625)),
aes(color = "normal distribution"), color = "bisque4", size = 1,  show.legend = FALSE) + 
geom_vline(aes(xintercept = mean(mns), color = 'sample mean'), size = 1, color = 'bisque3') +
geom_vline(aes(xintercept = 5, color = 'theoretical mean'), size = 1, linetype = 'dashed', color = 'bisque4')
```

